```{r,echo=FALSE}
opts_chunk$set(tidy=FALSE)
```

# Examples from lesson number 1 - Intro to R - presented by Rich FitzJohn  

## Vector types

Rich just told us that in R we have 5 different types of vector.  
```{r}
x <- TRUE       # x becomes TRUE
x <- 1          # x now has a value of 1 (overwritten the line above)
x <- c(1, TRUE) # x now contains two values of 1, because a vector always stores the same mode of elements
x
x <- c(1, TRUE, '1') #now all values are characters, because there is a underlying hierarchy. For instance, if there is one element of mode 'character' in your vector, then all elements will be treated as so
x
```

## Lists  
If you want to store the different modes of elements in one single object, you use lists.  
```{r}
x <- list(1, TRUE, "1")
x
```

## Querrying  
You can ask yourself about the nature of an unknown vector - suppose you did not know what x is:  
```{r}
storage.mode(x)
typeof(x)
class(x)
```

# Examples from lesson number 2 - Writing functions - presented by Rich FitzJohn  

## Playing with the basics  
When you write your function, you define it based on *arguments*  
```{r}
f <- function(something) {
    something
}
f(100)
```
The function `f()` is what we call the identity function, i.e., the function returns the exact input. In this particular case, the argument `something` took the value 100. You could instead define an object before, and then calling the function on that object, which should do exactly the same thing.  
```{r}
x  <-  100 #x now is a numeric vector of one element of value 100
f(x)
```
It returned exactly the same thing. In this case, `something` took the value of x, which was 100.  

## Downloading the gapminder data
We have to download a .csv file that we will use for the rest of our course. Right-click this [link][id] and save it to some new directory. Let's play with it now. 
```{r}
data  <-  read.csv('gapminder-FiveYearData.csv', header=TRUE, stringsAsFactors=FALSE)
head(data) #try the function tail() instead, what happened?

#subset the data for year 1982 only
data.1982  <-  data[data$year == 1982, ]

#let's create a function to compute the arithmetic average of a vector
average  <-  function(x) {
    sum(x) / length(x)
}

average(data.1982$gdpPercap)
#same as 
sum(data.1982$gdpPercap) / length(data.1982$gdpPercap)
```
Notice that repeating things may be problematic; first because it increases your chances of typos; second, if for some reason you change the name of your columns in `dat`, you will have to manually change all these names, which will be quite annoying.  

## Making a plot  
```{r, echo=FALSE}
linear.rescale  <-  function(x, range) {
    p  <-  (x - min(x)) / (max(x) - min(x))
    range[[1]] + p * (range[[2]] - range[[1]])
}

colour.by.category <- function(things, table) {
    unname(table[things])
}

add.trend.line  <-  function(x, y, d, col) {
    fit  <-  lm(d[[y]]~log10(d[[x]]))
    abline(fit, col=col, lwd=1.5, lty=2)
}

my.plot  <-  function(year, data, cols) {
    data.year  <-  data[data$year == year, ]
    col       <-  colour.by.category(data.year$continent, cols)
    cex       <-  linear.rescale(sqrt(data.year$pop), range=c(0.2,10))
    plot(lifeExp ~ gdpPercap, data.year, las=1, xlab='GDP per capta', ylab='Life expectancy', cex=cex, col='black', bg=col, pch=21, log="x", xlim=c(240,114000), ylim=c(20,85), main=unique(data.year$year))
    for(continent in unique(data.year$continent)) {
        add.trend.line("gdpPercap", "lifeExp",
                       d=data.year[data.year$continent == continent,],
                       col=cols[[continent]])
    }
}
```
```{r}
col.table  <-  c(Asia='tomato',
                 Europe='chocolate4',
                 Africa='dodgerblue2',
                 Americas='darkgoldenrod1',
                 Oceania='green4')

for (year in unique(data$year))
    my.plot(year, data, col.table)

```

[id]: https://raw.github.com/nicercode/gapminder/master/gapminder-FiveYearData.csv